---
title: "readr"
format: html
---



```{r}
library(tidyverse) # for functions beyond readr
library(gt) # to get the exibble package
library(readr)
```

Interesting fun fact I've learned while creating fake data for my data cleaning course:

You can chain multiple `write_*()` functions from the {readr} package.

All of these invisibly return the original data.

So you can just pass that invisible output to the next write function.

That way, you can save the same data set to a myriad of formats quick & easy.

Proably a rare use case but for me it was just the right thing to create a bunch of different file formats for teaching.



```{r}
exibble |> 
  write_csv(
    "O1_exibble_no_headers.csv", col_name = F
  ) |> 
  write_csv("02_exibble.csv") |> 
  write_csv2(
    "03_exibble_european.csv"
    ) |> 
  write_tsv(
  "04_exibble.tsv"
  ) |> 
  write_csv(
    "05_exibble_with_metadata_csv", 
    append = TRUE, 
    col_names = TRUE
  ) |> 
  write_delim(
    "12_weird_delimiter_text", delim = "|"
  )
  
  
```

